Fast Algorithms for Intersection of Non-matching Grids Using \plucker Coordinates.

### 2016-12-14 paper_intersections.tex

\begin{lemma}
\label{lemma_barycentric}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $\vc X$ on the $p$ and inside the triangle $S_2$. 
The barycentric coordinates of $\vc X$ on $S_2$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates, the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ has \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions and substituing for $\vc V_0-\vc V_2=\vc s_1$, we get for side $\vc s_1$
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\begin{eqnarray}
   \pi_p \odot \pi_{s_i} &=& -w_i \vc u \cdot (\vc s_1 \times \vc s_2), \label{eq:bary_centric_part1}\\
   \sum_{i=0}^{2} \pi_p \odot \pi_{s_i} &=& - \vc u \cdot (\vc s_1 \times \vc s_2). \label{eq:bary_centric_part2}
\end{eqnarray}
The result \eqref{eq:bary_centric} then follows directly from combination of \eqref{eq:bary_centric_part1} and \eqref{eq:bary_centric_part2}.
The point $\vc X$ is inside $S_2$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $\vc X$ on $S_2$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

@@

\begin{figure}[!htb]
    \centering    
    \includegraphics[width=0.9\textwidth]{\figpath vertex_faces.pdf}
    \caption{Possible cases processed in the function \emph{vertex faces}.
             Only the main features refered in text are denoted: tetrahedorn vertex $v$, edge $e$, face $f$ (stripes).}
    \label{fig:vertex_faces}
\end{figure}

The function in \algo{algo:vertex_faces} gets as a parameter IC $p$ at the vertex $v$ of $S_3$ 
which is a special case of a non-degenerate edge-triangle intersection. There are three
edges incident with the vertex $v$, results $s[i]$ of their intersections with $S_2$ may be one of: no IC, 
degenerate IC, positive IC and negative IC. Accordingly we call the edge: without intersection, degenerate, positive, or negative. 
We use these edge idicators to return generalized faces of $S_3$ preceding and succeeding $p$ on the polygons boundary assuming $p$ is at interior of $S_2$.
Possible cases are (see also \fig{fig:vertex_faces} a-e):
\begin{itemize}
 \item {\bf Single degenerate IC} (line \ref{line:one_degenerated}){\bf .} Let us denote $e$ the edge with degenerate IC 
 and $f$ the face between the other two edges.
 The other two (non-degenerate) edges may have either the opposite sign (the plane is cutting $S_3$, see \fig{fig:vertex_faces}, (a)) 
 or the same sign (the plane is touching $S_3$ at the edge $e$, see \fig{fig:vertex_faces}, (b)).
 In the first case, the call of edge faces for $e$ returns $(f_x,f)$ or $(f,f_x)$, then the vertex faces function returns $(e, f)$ or $(f, e)$, respectively.
 In the second case, we just return anything since there are at most two ICs in IP so the connection tables are not necessary. However, we add 
 current IC into $J$.

>>>>

\begin{lemma}
\label{lemma_barycentric}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $\vc X$ on the $p$ and inside the triangle $S_2$. 
The barycentric coordinates of $\vc X$ on $S_2$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates, the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ has \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions and substituting for $\vc V_0-\vc V_2=\vc s_1$, we get for side $\vc s_1$
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\begin{eqnarray}
   \pi_p \odot \pi_{s_i} &=& -w_i \vc u \cdot (\vc s_1 \times \vc s_2), \label{eq:bary_centric_part1}\\
   \sum_{i=0}^{2} \pi_p \odot \pi_{s_i} &=& - \vc u \cdot (\vc s_1 \times \vc s_2). \label{eq:bary_centric_part2}
\end{eqnarray}
The result \eqref{eq:bary_centric} then follows directly from combination of \eqref{eq:bary_centric_part1} and \eqref{eq:bary_centric_part2}.
The point $\vc X$ is inside $S_2$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $\vc X$ on $S_2$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

@@

\begin{figure}[!htb]
    \centering    
    \includegraphics[width=0.9\textwidth]{\figpath vertex_faces.pdf}
    \caption{Possible cases processed in the function \emph{vertex faces}.
             Only the main features referred in text are denoted: tetrahedron vertex $v$, edge $e$, face $f$ (stripes).}
    \label{fig:vertex_faces}
\end{figure}

The function in \algo{algo:vertex_faces} gets as a parameter IC $p$ at the vertex $v$ of $S_3$ 
which is a special case of a non-degenerate edge-triangle intersection. There are three
edges incident with the vertex $v$, results $s[i]$ of their intersections with $S_2$ may be one of: no IC, 
degenerate IC, positive IC and negative IC. Accordingly we call the edge: without intersection, degenerate, positive, or negative. 
We use these edge indicators to return generalized faces of $S_3$ preceding and succeeding $p$ on the polygons boundary assuming $p$ is at interior of $S_2$.
Possible cases are (see also \fig{fig:vertex_faces} a-e):
\begin{itemize}
 \item {\bf Single degenerate IC} (line \ref{line:one_degenerated}){\bf .} Let us denote $e$ the edge with degenerate IC 
 and $f$ the face between the other two edges.
 The other two (non-degenerate) edges may have either the opposite sign (the plane is cutting $S_3$, see \fig{fig:vertex_faces}, (a)) 
 or the same sign (the plane is touching $S_3$ at the edge $e$, see \fig{fig:vertex_faces}, (b)).
 In the first case, the call of edge faces for $e$ returns $(f_x,f)$ or $(f,f_x)$, then the vertex faces function returns $(e, f)$ or $(f, e)$, respectively.
 In the second case, we just return anything since there are at most two ICs in IP so the connection tables are not necessary. However, we add 
 current IC into $J$.
